Overview

### 2023-12-02 analysis/001_overview.Rmd
FInish publisher's league table

publisher_df <- left_join(jn_ind_jct_all, jn_ind_jct_oa) |>
  mutate(cr_year = as.character(cr_year)) |>
  left_join(ta_active) |>
  mutate(across(everything(), .fns = ~replace_na(.,0))) |>
  mutate(no_ta = oa - ta_oa)

## plot
publisher_plot_df <- publisher_df |> 
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value") |>
  mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )
  ) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           )) |>
  filter(!cr_year %in%  c("2017", "2023"))

@@

```{r publisher_summary_table}
publisher_article_total_table <- publisher_summary_df |>
  distinct(publisher, jn_all) |>
  group_by(publisher) |>
  summarise(publisher_total = sum(jn_all)) |>
  mutate(perc = publisher_total / sum(publisher_total),
         cum = cumsum(publisher_total) / sum(publisher_total))

@@

publisher_article_ta_oa_table <- publisher_summary_df |>
  filter(type == "ta_oa") |>
  group_by(publisher) |>
  summarise(ta_oa_total = sum(value)) |>
  mutate(ta_oa_perc = ta_oa_total / sum(ta_oa_total),
         ta_oa_cum = cumsum(ta_oa_total) / sum(ta_oa_total))

@@

kbl(publisher_league_tbl) %>%
  kable_classic() %>%
  add_header_above(c("Publisher" = 1, "Group 1" = 3, "Group 2" = 3, "Group 3" = 3))
```

>>>>

publisher_df <- left_join(jn_ind_jct_all, jn_ind_jct_oa) |>
  mutate(cr_year = as.character(cr_year)) |>
  left_join(ta_active) |>
  mutate(across(everything(), .fns = ~replace_na(.,0))) |>
   mutate(no_ta = oa - ta_oa) |>
    mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )
  ) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           )) |>
  filter(!cr_year %in%  c("2017", "2023")) |>
  # Journal transfers across publishers grouped in publisher
  select(-esac_publisher) |>
  distinct()

## plot
publisher_plot_df <- publisher_df |> 
  pivot_longer(cols = c(ta_oa, no_ta), names_to = "type", values_to = "value")

@@

```{r publisher_summary_table}
publisher_article_total_table <- publisher_df |>
  group_by(publisher) |>
  summarise(oa_journals = n_distinct(issn_l),
            publisher_total = sum(jn_all)) |>
  mutate(
    jn_perc = oa_journals / sum(oa_journals),
    jn_cum = cumsum(oa_journals) / sum(oa_journals),
    perc = publisher_total / sum(publisher_total),
    cum = cumsum(publisher_total) / sum(publisher_total)
  ) |>
  relocate(publisher_total, .after = jn_cum)

@@

publisher_article_ta_oa_table <- publisher_summary_df |>
  filter(type == "ta_oa") |>
  group_by(publisher) |>
  summarise(ta_oa_total = sum(value)) |>
  mutate(
    ta_oa_perc = ta_oa_total / sum(ta_oa_total),
    ta_oa_cum = cumsum(ta_oa_total) / sum(ta_oa_total)
  )

@@

publisher_league_tbl |>
  mutate(across(contains(c("perc", "cum")), ~ paste(round(., 3) * 100))) |>
  kbl(
    col.names = c(
      "Publisher",
      "Total",
      "%",
      "C%",
      "Total",
      "%",
      "C%",
      "Total",
      "%",
      "C%",
      "Total",
      "%",
      "C%"
    ),
    format.args = list(big.mark = ',')
  ) %>%
  kable_classic() %>%
  add_header_above(c(
    " " = 1,
    "Hybrid journals" = 3,
    "Articles" = 3,
    "OA articles" = 3,
    "TA OA articles" = 3
  ))
```
